\section{Day 5}

\subsection{Centrifugal Force Things}
Consider the acceleration in cylindrical coordinates.
\[ \vec{a} = \hat{s} (\ddot{s} - s \dot{\phi}) + \cdots. \]
$\ddot{s}$ is the familiar acceleration of the radial coordinate.
However, $-s\dot{\phi}$ is non-zero and keeps the rotating object
constrained to the circle it's moving along. However, in an accelerating
circular frame of reference, this term appears on the other side of the
equation.

\subsection{Terminal Velocity}
This occurs when there is some extra force $\left(\sum kv^{e_j}\right)
\hat{v}$. When there is some other (semi-)constant force, we have
$\vec{F} + \vec{F}_d = m \dot{\vec{v}}$, and $\dot{\vec{v}} \to 0$ for
some sufficiently large time $t$, we will reach some constant velocity,
which we term terminal velocity.

\subsection{Homework Remarks}
A useful trick for solving integrals of the form $\int_0^\infty e^{-t^2}
dt$ is to square and get the form $\iint e^{-(x^2 + y^2)} dx dy$ and
switch to polar coordinates to introduce $\int e^{-r^2} r dr d \theta$.

\subsection{Rocket Motion}
A rocket ejects exhaust out the back end with some speed $v_{e}$ relative
to the rocket. Suppose there is also some external force $\pvec{F}'$,
which is usually $0$, or gravity.

\begin{rem}
Here we introduce the actual form of Newton's second law.
\[ \vec{F} = \frac{d\vec{p}}{dt} \]
This equation never lies.
\end{rem}

Let the initial mass and velocity of the rocket be $m_0, v_0$,
respectively. After some timestep $\Delta t$, we have speed $v = v_0 +
\Delta v$ and eject $m_e$ exhaust mass at speed $v_e$, so our rocket
mass is now $m = m_0 - m_e$. Here we have the forms for momentum in this
case
\[ \begin{aligned}
p &= m_0 v_0 \\
p' &= mv + m_e(v_0 - v_e) \\
&= (m - m_e)(v_0 + \Delta v) + m_e(v_0 - v_e) \\
&= m_0 (v_0 + \Delta v) - m_e (v_e + \Delta v) \\
\end{aligned} \]
The difference yields
\[ \Delta p = m_0 \Delta v + \Delta m v_e + \Delta m \Delta v, \]
which implies
\[ \frac{dp}{dt} = m_0 \frac{dv}{dt} + \frac{dm}{dt} v_e. \]
Now we rearrange the force equation to get
\[ F' - \dot{m}v_e = m \dot{v}. \]
We also coin the term $-\dot{m}v_e$ as \textbf{thrust}.

When $F' = 0$, we have
\[ \begin{aligned}
-\dot{m} v_e &= m \dot{v} \\
-\frac{dm}{m} &= \frac{dv_e}{v_e} \\
\end{aligned} \]
After some integration, we get $v(m) = v_0 + v_e \ln \frac{m_0}{m}$.
