\section{September 25, 2015}

\subsection{Review}

Last time, we covered two big ideas.
\begin{enumerate}
\item Let $N$ be a normal subgroup of $G$. Then the cosets $G/N$ form a
group with the natural operation $(aN)(bN) = (ab)N$.
\item The First Isomorphism Theorem, or let $\varphi : G \to H$ be a
homomorphism. Then $\varphi$ induces a well defined isomorphism
$\overline{\varphi} : G / \ker \varphi \to \im \varphi$ given by
$\overline{\varphi}(a \ker \varphi) = \varphi(a)$ with a nice
commutative square.
\end{enumerate}

\subsection{Examples}
\begin{ex}
Let $\varphi : \RR \to \CC^*$ be given by $\theta \mapsto e^{2 \pi i
\theta}$. Then $\im \varphi = \varphi(\RR) = \lbrace z \in C^* : |z| = 1
\rbrace = S^1$ and $\ker \varphi = \lbrace \theta : e^{2 \pi i \theta} =
1 \rbrace = \ZZ$. The isomorphism theorem tells us that the isomorphism
$\overline{\varphi} : \RR / \ZZ \to S^1$ is well defined.
\end{ex}

\begin{ex}
We have the natural isomorphisms between $\ZZ / n\ZZ \iso \frac{1}{n}
\ZZ / \ZZ \iso \mu_n$.

Some elaboration... To get $\ZZ / n\ZZ \iso \frac{1}{n}\ZZ / \ZZ$, we
use the map $\varphi : \ZZ \to \frac{1}{n}\ZZ, k \mapsto \frac{k}{n}$
composed with the map $\pi : \frac{1}{n}\ZZ \to \frac{1}{n}\ZZ / \ZZ$.
Clearly $\pi \circ \varphi$ is a homomorphism and $\ker(\pi \circ
\varphi) = \lbrace k : \frac{k}{n} \in \ZZ \rbrace = n\ZZ$. Then the
first isomorphism theorem gets us that there is some unique and well
defined isomorphism $\overline{\pi \circ \varphi} : \ZZ / n\ZZ \to
\frac{1}{n}\ZZ / \ZZ, a + n\ZZ \mapsto \frac{a}{n} + \ZZ$.

We also have the silly isomorphism $\psi : \frac{1}{n}\ZZ \surj \mu_n$
by $\frac{k}{n} \mapsto e^{\frac{2 i \pi k}{n}}$ which has kernel $\ZZ$.
\end{ex}

\subsection{Products of groups}
Let $K, H$ be two groups. We can turn their cartesian product into a
group by doing things element wise, or
\[ (k_1, h_1) \cdot (k_2, h_2) = (k_1 k_2, h_1 h_2) \]
is a group.

We have some nice properties as well.
\begin{enumerate}
\item We have projection maps $\pi_K : K \times H \to K$ and $\pi_H : K
\times H \to H$.
\item For all pairs of homomorphisms $\varphi : G \to K, \psi : G \to
H, \exists ! \quad \varphi \times \psi : G \to K \times H$ a
homomorphism.
\end{enumerate}

\begin{ex}
$\ZZ / 6 \ZZ \iso \ZZ / 2 \ZZ \times \ZZ / 3 \ZZ$.
\end{ex}

\begin{proof}
We have the obvious
homomorphism $f : \ZZ \to \ZZ / 2 \ZZ \times \ZZ / 3 \ZZ$ given by $k
\mapsto ([k]_2, [k]_3)$. Combined with the fact that $\ker f = 6\ZZ$,
the first isomorphism theorem says that $\overline{f} : \ZZ / 6 \ZZ \to
\ZZ / 2 \ZZ \times \ZZ / 3 \ZZ$ given by $[k]_6 \mapsto ([k]_2, [k]_3)$
is a well defined isomorphism.
\end{proof}

\subsection{New Things?}
Nope. There was a quiz.
